Another DSL primer

### 2013-11-12 docs/fsml/paper.tex
Made the accept predicates a bit more regular

Let us define the syntax in terms of a context-free grammar. We use an
EGL grammar, where EGL stands for `extended grammar language' and can
be regarded as a variation on the Extended Backus Naur Form. EGL is an
ad-hoc grammar notation that is provided by the \slepro{} project; see
\S\ref{A:egl} for a self-description. In particular, EGL grammars can
be executed for the purpose of parsing, but we omit all such details
here. See \autoref{languages/fsml/cs.egl} for the grammar. It can be
assumed that the concrete textual syntax, as illustrated in
\autoref{languages/fsml/sample-turnstile.fsml}, can be parsed indeed
with the present grammar.

@@

Let us define the abstract syntax in terms of a signature that defines
all valid FSM terms. We use an ESL signature, where ESL stands for
`extended signature language' and can be regarded as (quite) a
variation on notations familiar from algebraic specification or (less)
a variation on type systems used in logic or functional
programming. ESL is an ad-hoc signature notation that is provided by
the \slepro{} project; see \S\ref{A:esl} for a syntax definition of
the signature notation. In particular, ESL grammars can be executed
for conformance checking to see whether a given term conforms to a
given signature, but we omit all such details here. See
\autoref{languages/fsml/as.esl} for the signature. The term in
\autoref{languages/fsml/sample-turnstile-fsml.term} conforms indeed to
the present signature.

@@

\autoref{languages/fsml/java/State.java}--\autoref{languages/fsml/java/Action.java}
shows (generated) Java enum types for the states, inputs, and
actions. Obviously, these sets are trivially obtainable from the
abstract syntactical representation of an FSM. In
\autoref{languages/fsml/to-java/state.pro}, this is illustrated for
the enum type for states. The predicate collects all (declared) state
ids and assembles an enum type, subject to a suitable abstract syntax
of Java; see \S\ref{A:java} for a summary of the (abstract) Java
syntax used by the code generator. Finally, the Java declaration is
pretty printed; see the use of \emph{ppJavaDecl/2} in the figure,
subject to a pretty printer for the Java language. The generator
predicates for the enum types for inputs and actions are quite
similar.

@@

The visualization in \autoref{F:turnstile} renders an FSM as a
graph. There are nodes and edges; both of them can be labeled; edges
are directed; nodes are of a certain shape (an ellipse here) and style
(regular or
filled). \autoref{languages/fsml/sample-turnstile-dgl.term} represents
represents the graph in the abstract (Prolog-based) syntax of a simple
graph description based on the DGL notation provided by the \slepro{}
project. DGL stands for `dot-based graph language'; see \S\ref{A:dgl}
for this notation. Nodes are quadruplets of the form (\m{Id},
\m{Label}, \m{Shape}, \m{Style}); edges are triplets of the form
(\m{FromId}, \m{ToId}, \m{OptionalLabel}).

>>>>

Let us define the syntax in terms of a context-free grammar. We use an
EGL grammar, where EGL stands for `extended grammar language' and can
be regarded as a variation on the Extended Backus Naur Form. EGL is an
ad-hoc grammar notation that is provided by the \slepro{} project; see
\ref{A:egl} for a self-description. In particular, EGL grammars can
be executed for the purpose of parsing, but we omit all such details
here. See \autoref{languages/fsml/cs.egl} for the grammar. It can be
assumed that the concrete textual syntax, as illustrated in
\autoref{languages/fsml/sample-turnstile.fsml}, can be parsed indeed
with the present grammar.

@@

Let us define the abstract syntax in terms of a signature that defines
all valid FSM terms. We use an ESL signature, where ESL stands for
`extended signature language' and can be regarded as (quite) a
variation on notations familiar from algebraic specification or (less)
a variation on type systems used in logic or functional
programming. ESL is an ad-hoc signature notation that is provided by
the \slepro{} project; see \ref{A:esl} for a syntax definition of
the signature notation. In particular, ESL grammars can be executed
for conformance checking to see whether a given term conforms to a
given signature, but we omit all such details here. See
\autoref{languages/fsml/as.esl} for the signature. The term in
\autoref{languages/fsml/sample-turnstile-fsml.term} conforms indeed to
the present signature.

@@

\autoref{languages/fsml/java/State.java}--\autoref{languages/fsml/java/Action.java}
shows (generated) Java enum types for the states, inputs, and
actions. Obviously, these sets are trivially obtainable from the
abstract syntactical representation of an FSM. In
\autoref{languages/fsml/to-java/state.pro}, this is illustrated for
the enum type for states. The predicate collects all (declared) state
ids and assembles an enum type, subject to a suitable abstract syntax
of Java; see \ref{A:java} for a summary of the (abstract) Java
syntax used by the code generator. Finally, the Java declaration is
pretty printed; see the use of \emph{ppJavaDecl/2} in the figure,
subject to a pretty printer for the Java language. The generator
predicates for the enum types for inputs and actions are quite
similar.

@@

The visualization in \autoref{F:turnstile} renders an FSM as a
graph. There are nodes and edges; both of them can be labeled; edges
are directed; nodes are of a certain shape (an ellipse here) and style
(regular or
filled). \autoref{languages/fsml/sample-turnstile-dgl.term} represents
represents the graph in the abstract (Prolog-based) syntax of a simple
graph description based on the DGL notation provided by the \slepro{}
project. DGL stands for `dot-based graph language'; see \ref{A:dgl}
for this notation. Nodes are quadruplets of the form (\m{Id},
\m{Label}, \m{Shape}, \m{Style}); edges are triplets of the form
(\m{FromId}, \m{ToId}, \m{OptionalLabel}).
